The Unicode Cookbook \linebreak for Linguists

### 2018-06-05 book/unicode.tex
deals with paperhivers comments for chapter 7 Orthography profiles

% latex table generated in R 3.4.3 by xtable 1.8-2 package
% Mon Mar 12 17:57:11 2018
\begin{table}[H]
\centering
\begingroup\scriptsize
\begin{tabular}{llll}
  \toprule
Gr. & Freq. & Codepoint & Unicode Name \\ 
  \midrule
Ú & 1 & U+00DA & LATIN CAPITAL LETTER U WITH ACUTE \\ 
  Ú & 1 & U+0055, U+0301 & LATIN CAPITAL LETTER U, COMBINING ACUTE ACCENT \\ 
  Ù & 1 & U+00D9 & LATIN CAPITAL LETTER U WITH GRAVE \\ 
  Ù & 1 & U+0055, U+0300 & LATIN CAPITAL LETTER U, COMBINING GRAVE ACCENT \\ 
  Û & 1 & U+00DB & LATIN CAPITAL LETTER U WITH CIRCUMFLEX \\ 
  Û & 1 & U+0055, U+0302 & LATIN CAPITAL LETTER U, COMBINING CIRCUMFLEX ACCENT \\ 
   \bottomrule
\end{tabular}
\endgroup
\caption{Profile 1 (default settings, splitting grapheme clusters)} 
\label{tab:profile1}
\end{table}

@@

% latex table generated in R 3.4.3 by xtable 1.8-2 package
% Mon Mar 12 17:57:11 2018
\begin{table}[H]
\centering
\begingroup\scriptsize
\begin{tabular}{llll}
  \toprule
Grapheme & Frequency & Codepoint & Unicode Name \\ 
  \midrule
́ & 1 & U+0301 & COMBINING ACUTE ACCENT \\ 
  ̀ & 1 & U+0300 & COMBINING GRAVE ACCENT \\ 
  ̂ & 1 & U+0302 & COMBINING CIRCUMFLEX ACCENT \\ 
  U & 3 & U+0055 & LATIN CAPITAL LETTER U \\ 
  Ú & 1 & U+00DA & LATIN CAPITAL LETTER U WITH ACUTE \\ 
  Ù & 1 & U+00D9 & LATIN CAPITAL LETTER U WITH GRAVE \\ 
  Û & 1 & U+00DB & LATIN CAPITAL LETTER U WITH CIRCUMFLEX \\ 
   \bottomrule
\end{tabular}
\endgroup
\caption{Profile 2 (splitting by code points)} 
\label{tab:profile2}
\end{table}

Some characters look identical, although they are encoded differently.
Unicode offers different ways of normalization (see
Section~\ref{pitfall-canonical-equivalence}), which can be invoked here as well
using the option \texttt{normalize}. NFC-normalization turns everything into the
precomposed characters, while NFD-normalization separates everything into base
characters with combining diacritics. Splitting by code points (i.e. \texttt{sep~=~""}) shows the results of these two normalizations.

@@

% latex table generated in R 3.4.3 by xtable 1.8-2 package
% Mon Mar 12 17:57:11 2018
\begin{table}[H]
\centering
\begingroup\scriptsize
\begin{tabular}{llll}
  \toprule
Grapheme & Frequency & Codepoint & Unicode Name \\ 
  \midrule
Ú & 2 & U+00DA & LATIN CAPITAL LETTER U WITH ACUTE \\ 
  Ù & 2 & U+00D9 & LATIN CAPITAL LETTER U WITH GRAVE \\ 
  Û & 2 & U+00DB & LATIN CAPITAL LETTER U WITH CIRCUMFLEX \\ 
   \bottomrule
\end{tabular}
\endgroup
\caption{Profile 3 (splitting by NFC code points)} 
\label{tab:profile3}
\end{table}

% latex table generated in R 3.4.3 by xtable 1.8-2 package
% Mon Mar 12 17:57:11 2018
\begin{table}[H]
\centering
\begingroup\scriptsize
\begin{tabular}{llll}
  \toprule
Grapheme & Frequency & Codepoint & Unicode Name \\ 
  \midrule
́ & 2 & U+0301 & COMBINING ACUTE ACCENT \\ 
  ̀ & 2 & U+0300 & COMBINING GRAVE ACCENT \\ 
  ̂ & 2 & U+0302 & COMBINING CIRCUMFLEX ACCENT \\ 
  U & 6 & U+0055 & LATIN CAPITAL LETTER U \\ 
   \bottomrule
\end{tabular}
\endgroup
\caption{Profile 4 (splitting by NFD code points)} 
\label{tab:profile4}
\end{table}

@@

% latex table generated in R 3.4.3 by xtable 1.8-2 package
% Mon Mar 12 17:57:11 2018
\begin{table}[H]
\centering
\begingroup\scriptsize
\begin{tabular}{llll}
  \toprule
Gr. & Freq. & Codepoint & Unicode Name \\ 
  \midrule
Ú & 2 & U+0055, U+0301 & LATIN CAPITAL LETTER U, COMBINING ACUTE ACCENT \\ 
  Ù & 2 & U+0055, U+0300 & LATIN CAPITAL LETTER U, COMBINING GRAVE ACCENT \\ 
  Û & 2 & U+0055, U+0302 & LATIN CAPITAL LETTER U, COMBINING CIRCUMFLEX ACCENT \\ 
   \bottomrule
\end{tabular}
\endgroup
\caption{Profile 5 (splitting by graphemes after NFD)} 
\label{tab:profile5}
\end{table}

% latex table generated in R 3.4.3 by xtable 1.8-2 package
% Mon Mar 12 17:57:11 2018
\begin{table}[H]
\centering
\begingroup\scriptsize
\begin{tabular}{llll}
  \toprule
Gr. & Freq. & Codepoint & Unicode Name \\ 
  \midrule
Ú & 2 & U+00DA & LATIN CAPITAL LETTER U WITH ACUTE \\ 
  Ù & 2 & U+00D9 & LATIN CAPITAL LETTER U WITH GRAVE \\ 
  Û & 2 & U+00DB & LATIN CAPITAL LETTER U WITH CIRCUMFLEX \\ 
   \bottomrule
\end{tabular}
\endgroup
\caption{Profile 6 (splitting by graphemes after NFC)} 
\label{tab:profile6}
\end{table}

@@

% latex table generated in R 3.4.3 by xtable 1.8-2 package
% Mon Mar 12 17:57:11 2018
\begin{table}[htb]
\centering
\begingroup\scriptsize
\begin{tabular}{llll}
  \toprule
Grapheme & Frequency & Codepoint & UnicodeName \\ 
  \midrule
sh &  &  &  \\ 
  ch &  &  &  \\ 
  sch &  &  &  \\ 
  ie &  &  &  \\ 
  oo &  &  &  \\ 
  a & 2 & U+0061 & LATIN SMALL LETTER A \\ 
  c & 2 & U+0063 & LATIN SMALL LETTER C \\ 
  e & 3 & U+0065 & LATIN SMALL LETTER E \\ 
  f & 1 & U+0066 & LATIN SMALL LETTER F \\ 
  h & 4 & U+0068 & LATIN SMALL LETTER H \\ 
  i & 3 & U+0069 & LATIN SMALL LETTER I \\ 
  m & 4 & U+006D & LATIN SMALL LETTER M \\ 
  o & 2 & U+006F & LATIN SMALL LETTER O \\ 
  p & 2 & U+0070 & LATIN SMALL LETTER P \\ 
  s & 4 & U+0073 & LATIN SMALL LETTER S \\ 
   \bottomrule
\end{tabular}
\endgroup
\caption{Manually edited profile skeleton} 
\label{tab:profile_skeleton1}
\end{table}

@@

% latex table generated in R 3.4.3 by xtable 1.8-2 package
% Mon Mar 12 17:57:11 2018
\begin{table}[htb]
\centering
\begingroup\scriptsize
\begin{tabular}{lllll}
  \toprule
Left & Grapheme & Frequency & Codepoint & UnicodeName \\ 
  \midrule
mi & s &  &  &  \\ 
   & sh &  &  &  \\ 
   & ch &  &  &  \\ 
   & sch &  &  &  \\ 
   & ie &  &  &  \\ 
   & oo &  &  &  \\ 
   & a & 2 & U+0061 & LATIN SMALL LETTER A \\ 
   & c & 2 & U+0063 & LATIN SMALL LETTER C \\ 
   & e & 3 & U+0065 & LATIN SMALL LETTER E \\ 
   & f & 1 & U+0066 & LATIN SMALL LETTER F \\ 
   & h & 4 & U+0068 & LATIN SMALL LETTER H \\ 
   & i & 3 & U+0069 & LATIN SMALL LETTER I \\ 
   & m & 4 & U+006D & LATIN SMALL LETTER M \\ 
   & o & 2 & U+006F & LATIN SMALL LETTER O \\ 
   & p & 2 & U+0070 & LATIN SMALL LETTER P \\ 
   & s & 4 & U+0073 & LATIN SMALL LETTER S \\ 
   \bottomrule
\end{tabular}
\endgroup
\caption{Orthography profile with contextual specification for <s>} 
\label{tab:profile_skeleton2}
\end{table}

@@

% latex table generated in R 3.4.3 by xtable 1.8-2 package
% Mon Mar 12 17:57:11 2018
\begin{table}[htb]
\centering
\begingroup\scriptsize
\begin{tabular}{lllll}
  \toprule
Left & Grapheme & Right & Class & Replacement \\ 
  \midrule
 & a &  &  & a \\ 
   & c &  &  & c \\ 
   & e &  &  & e \\ 
   & i &  &  & i \\ 
   & n &  &  & n \\ 
   \bottomrule
\end{tabular}
\endgroup
\caption{Orthography profile with empty columns for editing contexts} 
\label{tab:profile_editing_1}
\end{table}

@@

% latex table generated in R 3.4.3 by xtable 1.8-2 package
% Mon Mar 12 17:57:11 2018
\begin{table}[htb]
\centering
\begingroup\scriptsize
\begin{tabular}{lllll}
  \toprule
Left & Grapheme & Right & Class & IPA \\ 
  \midrule
 & c & [ie] &  & tʃ \\ 
   & a &  &  & a \\ 
   & n &  &  & n \\ 
   & c &  &  & k \\ 
   & e &  &  & e \\ 
   & i &  &  & i \\ 
   \bottomrule
\end{tabular}
\endgroup
\caption{Orthography profile with regex as context} 
\label{tab:profile_editing_2}
\end{table}

@@

% latex table generated in R 3.4.3 by xtable 1.8-2 package
% Mon Mar 12 17:57:11 2018
\begin{table}[htb]
\centering
\begingroup\scriptsize
\begin{tabular}{lllll}
  \toprule
Left & Grapheme & Right & Class & IPA \\ 
  \midrule
 & c & Vfront &  & tʃ \\ 
   & a &  &  & a \\ 
   & n &  &  & n \\ 
   & c &  &  & k \\ 
   & e &  & Vfront & e \\ 
   & i &  & Vfront & i \\ 
   \bottomrule
\end{tabular}
\endgroup
\caption{Orthography profile with Class as context} 
\label{tab:profile_editing_3}
\end{table}

>>>>

% latex table generated in R 3.5.0 by xtable 1.8-2 package
% Mon Jun  4 10:30:43 2018
\begin{table}[H]
\centering
\begingroup\scriptsize
\begin{tabular}{llll}
  \toprule
Gr. & Freq. & Codepoint & Unicode Name \\ 
  \midrule
Ú & 1 & U+00DA & LATIN CAPITAL LETTER U WITH ACUTE \\ 
  Ú & 1 & U+0055, U+0301 & LATIN CAPITAL LETTER U, COMBINING ACUTE ACCENT \\ 
  Ù & 1 & U+00D9 & LATIN CAPITAL LETTER U WITH GRAVE \\ 
  Ù & 1 & U+0055, U+0300 & LATIN CAPITAL LETTER U, COMBINING GRAVE ACCENT \\ 
  Û & 1 & U+00DB & LATIN CAPITAL LETTER U WITH CIRCUMFLEX \\ 
  Û & 1 & U+0055, U+0302 & LATIN CAPITAL LETTER U, COMBINING CIRCUMFLEX ACCENT \\ 
   \bottomrule
\end{tabular}
\endgroup
\caption{Profile 1 (default settings, splitting grapheme clusters)} 
\label{tab:profile1}
\end{table}

@@

% latex table generated in R 3.5.0 by xtable 1.8-2 package
% Mon Jun  4 10:30:43 2018
\begin{table}[H]
\centering
\begingroup\scriptsize
\begin{tabular}{llll}
  \toprule
Grapheme & Frequency & Codepoint & Unicode Name \\ 
  \midrule
́ & 1 & U+0301 & COMBINING ACUTE ACCENT \\ 
  ̀ & 1 & U+0300 & COMBINING GRAVE ACCENT \\ 
  ̂ & 1 & U+0302 & COMBINING CIRCUMFLEX ACCENT \\ 
  U & 3 & U+0055 & LATIN CAPITAL LETTER U \\ 
  Ú & 1 & U+00DA & LATIN CAPITAL LETTER U WITH ACUTE \\ 
  Ù & 1 & U+00D9 & LATIN CAPITAL LETTER U WITH GRAVE \\ 
  Û & 1 & U+00DB & LATIN CAPITAL LETTER U WITH CIRCUMFLEX \\ 
   \bottomrule
\end{tabular}
\endgroup
\caption{Profile 2 (splitting by code points)} 
\label{tab:profile2}
\end{table}

Some characters look identical, although they are encoded differently.
Unicode offers different ways of normalization (see
Section~\ref{pitfall-canonical-equivalence}), which can be invoked here as well
using the option \texttt{normalize}. NFC normalization turns everything into the
precomposed characters, while NFD normalization separates everything into base
characters with combining diacritics. Splitting by code points (i.e. \texttt{sep~=~""}) shows the results of these two normalizations.

@@

% latex table generated in R 3.5.0 by xtable 1.8-2 package
% Mon Jun  4 10:30:43 2018
\begin{table}[H]
\centering
\begingroup\scriptsize
\begin{tabular}{llll}
  \toprule
Grapheme & Frequency & Codepoint & Unicode Name \\ 
  \midrule
Ú & 2 & U+00DA & LATIN CAPITAL LETTER U WITH ACUTE \\ 
  Ù & 2 & U+00D9 & LATIN CAPITAL LETTER U WITH GRAVE \\ 
  Û & 2 & U+00DB & LATIN CAPITAL LETTER U WITH CIRCUMFLEX \\ 
   \bottomrule
\end{tabular}
\endgroup
\caption{Profile 3 (splitting by NFC code points)} 
\label{tab:profile3}
\end{table}

% latex table generated in R 3.5.0 by xtable 1.8-2 package
% Mon Jun  4 10:30:43 2018
\begin{table}[H]
\centering
\begingroup\scriptsize
\begin{tabular}{llll}
  \toprule
Grapheme & Frequency & Codepoint & Unicode Name \\ 
  \midrule
́ & 2 & U+0301 & COMBINING ACUTE ACCENT \\ 
  ̀ & 2 & U+0300 & COMBINING GRAVE ACCENT \\ 
  ̂ & 2 & U+0302 & COMBINING CIRCUMFLEX ACCENT \\ 
  U & 6 & U+0055 & LATIN CAPITAL LETTER U \\ 
   \bottomrule
\end{tabular}
\endgroup
\caption{Profile 4 (splitting by NFD code points)} 
\label{tab:profile4}
\end{table}

@@

% latex table generated in R 3.5.0 by xtable 1.8-2 package
% Mon Jun  4 10:30:43 2018
\begin{table}[H]
\centering
\begingroup\scriptsize
\begin{tabular}{llll}
  \toprule
Gr. & Freq. & Codepoint & Unicode Name \\ 
  \midrule
Ú & 2 & U+0055, U+0301 & LATIN CAPITAL LETTER U, COMBINING ACUTE ACCENT \\ 
  Ù & 2 & U+0055, U+0300 & LATIN CAPITAL LETTER U, COMBINING GRAVE ACCENT \\ 
  Û & 2 & U+0055, U+0302 & LATIN CAPITAL LETTER U, COMBINING CIRCUMFLEX ACCENT \\ 
   \bottomrule
\end{tabular}
\endgroup
\caption{Profile 5 (splitting by graphemes after NFD)} 
\label{tab:profile5}
\end{table}

% latex table generated in R 3.5.0 by xtable 1.8-2 package
% Mon Jun  4 10:30:43 2018
\begin{table}[H]
\centering
\begingroup\scriptsize
\begin{tabular}{llll}
  \toprule
Gr. & Freq. & Codepoint & Unicode Name \\ 
  \midrule
Ú & 2 & U+00DA & LATIN CAPITAL LETTER U WITH ACUTE \\ 
  Ù & 2 & U+00D9 & LATIN CAPITAL LETTER U WITH GRAVE \\ 
  Û & 2 & U+00DB & LATIN CAPITAL LETTER U WITH CIRCUMFLEX \\ 
   \bottomrule
\end{tabular}
\endgroup
\caption{Profile 6 (splitting by graphemes after NFC)} 
\label{tab:profile6}
\end{table}

@@

% latex table generated in R 3.5.0 by xtable 1.8-2 package
% Mon Jun  4 10:30:43 2018
\begin{table}[htb]
\centering
\begingroup\scriptsize
\begin{tabular}{llll}
  \toprule
Grapheme & Frequency & Codepoint & UnicodeName \\ 
  \midrule
sh &  &  &  \\ 
  ch &  &  &  \\ 
  sch &  &  &  \\ 
  ie &  &  &  \\ 
  oo &  &  &  \\ 
  a & 2 & U+0061 & LATIN SMALL LETTER A \\ 
  c & 2 & U+0063 & LATIN SMALL LETTER C \\ 
  e & 3 & U+0065 & LATIN SMALL LETTER E \\ 
  f & 1 & U+0066 & LATIN SMALL LETTER F \\ 
  h & 4 & U+0068 & LATIN SMALL LETTER H \\ 
  i & 3 & U+0069 & LATIN SMALL LETTER I \\ 
  m & 4 & U+006D & LATIN SMALL LETTER M \\ 
  o & 2 & U+006F & LATIN SMALL LETTER O \\ 
  p & 2 & U+0070 & LATIN SMALL LETTER P \\ 
  s & 4 & U+0073 & LATIN SMALL LETTER S \\ 
   \bottomrule
\end{tabular}
\endgroup
\caption{Manually edited profile skeleton} 
\label{tab:profile_skeleton1}
\end{table}

@@

% latex table generated in R 3.5.0 by xtable 1.8-2 package
% Mon Jun  4 10:30:43 2018
\begin{table}[htb]
\centering
\begingroup\scriptsize
\begin{tabular}{lllll}
  \toprule
Left & Grapheme & Frequency & Codepoint & UnicodeName \\ 
  \midrule
mi & s &  &  &  \\ 
   & sh &  &  &  \\ 
   & ch &  &  &  \\ 
   & sch &  &  &  \\ 
   & ie &  &  &  \\ 
   & oo &  &  &  \\ 
   & a & 2 & U+0061 & LATIN SMALL LETTER A \\ 
   & c & 2 & U+0063 & LATIN SMALL LETTER C \\ 
   & e & 3 & U+0065 & LATIN SMALL LETTER E \\ 
   & f & 1 & U+0066 & LATIN SMALL LETTER F \\ 
   & h & 4 & U+0068 & LATIN SMALL LETTER H \\ 
   & i & 3 & U+0069 & LATIN SMALL LETTER I \\ 
   & m & 4 & U+006D & LATIN SMALL LETTER M \\ 
   & o & 2 & U+006F & LATIN SMALL LETTER O \\ 
   & p & 2 & U+0070 & LATIN SMALL LETTER P \\ 
   & s & 4 & U+0073 & LATIN SMALL LETTER S \\ 
   \bottomrule
\end{tabular}
\endgroup
\caption{Orthography profile with contextual specification for <s>} 
\label{tab:profile_skeleton2}
\end{table}

@@

% latex table generated in R 3.5.0 by xtable 1.8-2 package
% Mon Jun  4 10:30:43 2018
\begin{table}[htb]
\centering
\begingroup\scriptsize
\begin{tabular}{lllll}
  \toprule
Left & Grapheme & Right & Class & Replacement \\ 
  \midrule
 & a &  &  & a \\ 
   & c &  &  & c \\ 
   & e &  &  & e \\ 
   & i &  &  & i \\ 
   & n &  &  & n \\ 
   \bottomrule
\end{tabular}
\endgroup
\caption{Orthography profile with empty columns for editing contexts} 
\label{tab:profile_editing_1}
\end{table}

@@

% latex table generated in R 3.5.0 by xtable 1.8-2 package
% Mon Jun  4 10:30:43 2018
\begin{table}[htb]
\centering
\begingroup\scriptsize
\begin{tabular}{lllll}
  \toprule
Left & Grapheme & Right & Class & IPA \\ 
  \midrule
 & c & [ie] &  & tʃ \\ 
   & a &  &  & a \\ 
   & n &  &  & n \\ 
   & c &  &  & k \\ 
   & e &  &  & e \\ 
   & i &  &  & i \\ 
   \bottomrule
\end{tabular}
\endgroup
\caption{Orthography profile with regex as context} 
\label{tab:profile_editing_2}
\end{table}

@@

% latex table generated in R 3.5.0 by xtable 1.8-2 package
% Mon Jun  4 10:30:44 2018
\begin{table}[htb]
\centering
\begingroup\scriptsize
\begin{tabular}{lllll}
  \toprule
Left & Grapheme & Right & Class & IPA \\ 
  \midrule
 & c & Vfront &  & tʃ \\ 
   & a &  &  & a \\ 
   & n &  &  & n \\ 
   & c &  &  & k \\ 
   & e &  & Vfront & e \\ 
   & i &  & Vfront & i \\ 
   \bottomrule
\end{tabular}
\endgroup
\caption{Orthography profile with Class as context} 
\label{tab:profile_editing_3}
\end{table}
